---
title: "STEP-IN — Intervention Message Decision Logic"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{STEP-IN — Intervention Message Decision Logic}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
```

```{r packages, message=FALSE, warning=FALSE, include=FALSE}
library(stepinWearable)
library(dplyr)
library(lubridate)
```

## Summary of Intervention Logic

This article documents how the intervention module decides **which
message** template to send and **how to set or adjust** the targets: X
(steps/day), Y (minutes/day), and Z (target cadence: 80/90/100/110/120 steps/min
for Y).

### Analysis Windows

-   **Current Window:** The date range selected by the user (if exactly
    14 days are available, all are used; otherwise, the default is the
    last 14 days).

-   **Previous Window:** The equivalent number of days immediately
    preceding the start of the Current Window.

### Key Performance Indicators (KPIs)

KPIs are calculated using the `kpis()` function over the daily summary
(`daily_summary()`):

-   `median_steps_day`: **Median** daily steps. (Updated from 'average'
    to 'median' for consistency with code)

-   `median_steps_80plus`, `median_steps_90plus`,
    `median_steps_100plus`, `median_steps_110plus`, `median_steps_120plus`: **Median** **minutes** per day at ≥80, ≥90, ≥100, ≥110 and ≥120 steps/min, respectively.

### Compliance Rules

-   **Step Compliance** (`steps_ok`): The current window's performance
    must show an improvement of proposed target (X) with an allowance of
    1% (e.g., if target was to increment 5%, an increment of at least 4%
    is considered valid), compared to the previous window: \$\$
    \\texttt{median_steps_day_cur} \\ge 1.04 \\times
    \\texttt{median_steps_day_prev} \$\$

-   **Minute Compliance** (`mins_ok`, starting from phase m4): The
    current window's performance must meet or exceed the last target
    minutes (Y) *minus 1* at the target cadence (Z): \$\$
    \\texttt{median_minutes_at_Z_cur} \\ge (Y - 1) \$\$

## Intervention Phases (State-Based)

The system gets the phase of intervention based on the saved state and
files previously processed for the same participant. The researcher is
allowed to change the phase of the intervention by selecting the t-index
of the file being processed.

Each `t` index represents a consecutive **2-week review period** (i.e.,
the 14-day window used to compute KPIs and generate the message for that cycle).
Importantly, **t = 0 is already part of the intervention workflow**, because it
corresponds to the first 2-week recording that is processed to generate the
first step target (X). Therefore, the onset of cadence goals occurs at `t = 5`
(i.e., after ~12 weeks / ~3 months have elapsed).

-   t = 0 → **`post_basal`**: The very first message (following
    baseline).

-   t: 1−4 → **`m1_3`**: Intervention months 1–3 (focused only on steps
    accumulation).

-   t = 5 → **`init_m4`**: Initialization of intensity phase (introduces
    Y min at a Z cadence target).

-   t ≥6 → **`m4_9`**: Intervention months 4–9 (steps + intensity with
    potential for cadence scaling).

### Phase I (post_basal)

First, the participant wears a Fitbit for two weeks, and this recording
acts as the baseline measure. Once the first 14 days have been recorded,
the user is ready to propose the first target.

This first target is focused on steps accumulation, where the steps
target (X) proposed is 5% above the median steps per day accumulated
during the baseline 14-day window recording: \$\$
\\texttt{median_steps_day} \* 1.05 \$\$

> In phase I, the user is not allowed to manually change the proposed
> target.

### Phase II (m1_3)

From there, every two weeks, the user will download a new recording from
the Fitbit. The first 12 wweeks (t-indices from 0 to 4) are focused on
increasing the volume of steps per day (X). After each recording is
processed, the user can generate a message to propose the next target,
which is:

-   If `median_steps_day` ≥ X\*, next X is `median_steps_day` \* 1.05

-   If `median_steps_day` \< X\*, next X is equal to previous X

> \*Note: there is an allowance of 1% to decide that the participant met
> the X target.
>
> Users are allowed to manually adjust the multiplication factor to set
> X in this phase.

### Phase III (init_m4)

In month 4, a new phase starts with the introduction of the minutes per
day (Y) walking at a given cadence (Z). To decide the initial target,
the process would explore the current 14-day window and set the target
as follows:

-   If `median_steps_110plus` ≥ 45\*, the target is determined by rounding 
    `median_steps_120plus` to the nearest multiple of 5 and then adding 5
    minutes, at a cadence of 120 steps/min (e.g., if participant accumulates
    a median of 48 min/day at 110 steps/minute and 28 minutes at 120 steps/minute,
    then the target would be Y = 35, Z = 120).

-   Else, if `median_steps_100plus` ≥ 45\*, the target is determined by rounding
    `median_steps_110plus` to the nearest multiple of 5 and then adding 5 minutes,
    at a cadence of 110 steps/min.

-   Else, if`median_steps_100plus` ≥ 5\* or `median_steps_90plus` ≥ 15\*,
    the target is determined by rounding `median_steps_100plus` to the nearest
    multiple of 5 and then adding 5 minutes, at a cadence of 100 steps/min.

-   Else, if `median_steps_90plus` ≥ 5\* or `median_steps_80plus` ≥ 15\*,
    the target is determined by rounding `median_steps_90plus` to the nearest
    multiple of 5 and then adding 5 minutes, at a cadence of 90 steps/min.

-   Else, the target is determined by rounding `median_steps_80plus` to the
    nearest multiple of 5 and then adding 5 minutes, at a cadence of 80 steps/min.
    
Whenever a Y target is (re)computed, it is derived from the participant’s
current median minutes at the target cadence (i.e., `median_steps_{Z}plus`).

> \*Note: there is an allowance of 1 minute to decide that the
> participant met the Y/Z target.
>
> Users are not allowed to manually adjust Y/Z target in this phase.

### Phase IV (m4_9)

Definition of the cadence progression logic (Y/Z):

- **Duration target (Y):** When (re)computed, Y is derived from the
participant’s current median minutes at the target cadence
(i.e., `median_steps_{Z}plus`), rounded to the nearest multiple of 5 and then
adding 5 minutes, until the escalation threshold for the current cadence is reached.

- **Cadence progression (Z):** When the escalation threshold is reached at the
current cadence, Z increases to the next level. After escalation, Y is computed
from the participant’s median minutes at the **new** cadence.

- **Escalation thresholds:** Z increases when the participant accumulates at
least **15 min/day** at Z = 80 or 90, and at least **45 min/day** at Z = 100 or
110.

Therefore, the cadence target (Z) is scaled up if the participant successfully
met the previous target:

-   If Z = 80 and `median_steps_80plus` ≥ 15 minutes → Next Z = 90
    steps/min and Y = the next integer multiple of 5 to `median_steps_90plus`.

-   If Z = 90 and `median_steps_90plus` ≥ 15 minutes → Next Z = 100
    steps/min and Y = the next integer multiple of 5 to `median_steps_100plus`.
    
-   If Z = 100 and `median_steps_100plus` ≥ 45 minutes → Next Z = 110
    steps/min and Y = the next integer multiple of 5 to `median_steps_110plus`.
    
-   If Z = 110 and `median_steps_110plus` ≥ 45 minutes → Next Z = 120
    steps/min and Y = the next integer multiple of 5 to `median_steps_120plus`.

Importantly, when cadence (Z) increases, Y is computed from the participant’s
current median minutes at the new target cadence.

Otherwise, progression of Y and Z follows the same logic as in Phase
III, considering that if participant meets the previous target, a new
target is proposed by rounding the current median to the nearest multiple of 5
and the adding 5 minutes, and otherwise, the previous target is replicated.

> Users are allowed to manually adjust the X, and the Y/Z targets in
> this phase.

### Handling Missing Data

If no data (missing rows or all zero values) is present in the **last 3
days of the current window**, the **`nodata3`** message is displayed,
and the saved targets are **not updated**.

------------------------------------------------------------------------

## Key Functions

-   `daily_summary(df)`: Converts the minute-level series (`timestamp`,
    `steps`) to a daily summary, including `steps_day`, `steps_80plus`,
    `steps_90plus`, `steps_100plus`, `steps_110plus`, and `steps_120plus`.

-   `select_windows(dsum, start_date, end_date)`: Constructs the
    **Current Window** (start–end) and the **Previous Window** (same
    number of preceding days).

-   `kpis(win)`: Calculates KPIs for a given window.

-   `decide_message(state, cur_k, prev_k, nombre)`: Applies the
    phase-based rules and returns:

    -   `key` (chosen template), `text` (rendered message), `next_X`,
        `next_Y`, `next_Z`, and `consecutive_fails`.

### Consecutive Failure Check

The decision engine tracks `consecutive_fails`. If a participant reaches
≥2 consecutive failures at meeting the proposed target, the message
generation will automatically include a **supportive paragraph** to
encourage compliance, regardless of the chosen message key.

------------------------------------------------------------------------

## Reproducible Example with Synthetic Data

We generate a minute-by-minute series over 20 days with simulated
activity patterns to demonstrate calculation of minutes at varying
cadences. We then pass this data through the core logic.

```{r simulate-minute-data}
set.seed(123)

# --- utility to simulate "walked" minutes at a simulated cadence ---
simulate_day <- function(date, base_steps = 6000) {
  # 1440 minutes/day
  minutes <- seq.POSIXt(
    as.POSIXct(paste(date, "00:00:00"), tz = "UTC"),
    by = "min", length.out = 1440
  )
  
  # Walking blocks: 2-4 blocks of 20-40 min, with cadence 70-110 steps/min
  n_blocks <- sample(2:4, 1)
  steps <- rep(0L, 1440)
  for (b in seq_len(n_blocks)) {
    len <- sample(20:40, 1)
    start <- sample(60:(1440 - len - 60), 1)
    cadence <- sample(70:110, 1)    # steps per minute
    steps[start:(start + len)] <- cadence
  }
  
  # Light noise/standing (0-10 steps)
  steps <- steps + rpois(1440, lambda = 1)
  
  tibble::tibble(
    timestamp = minutes,
    steps = as.integer(steps)
  )
}

# 20 synthetic days
dates <- seq.Date(Sys.Date() - 19, Sys.Date(), by = "day")
df_minute <- dplyr::bind_rows(lapply(dates, simulate_day))

# --- add simulated heart rate (per-minute) ---
df_minute <- df_minute |>
  dplyr::mutate(
    heart_rate = dplyr::case_when(
      steps == 0L ~ round(rnorm(dplyr::n(), mean = 65, sd = 5)),     # resting
      steps < 50L ~ round(rnorm(dplyr::n(), mean = 80, sd = 8)),     # light
      TRUE        ~ round(rnorm(dplyr::n(), mean = 110, sd = 10))    # walking
    ),
    heart_rate = pmin(pmax(heart_rate, 50), 180)  # clamp to plausible range
  )
```

### Summarize to daily data:

```{r daily-sum}
dsum <- stepinWearable:::daily_summary(df = df_minute)
head(dsum)
```

### Select a 14-day window with select_windows():

```{r windows}
end_date   <- max(dsum$date)
start_date <- end_date - 13  # 14 days
wins <- stepinWearable:::select_windows(
  dsum, 
  start_date = start_date, 
  end_date = end_date)

curk  <- stepinWearable:::kpis(wins$current)
prevk <- if (nrow(wins$previous) > 0) stepinWearable:::kpis(wins$previous) else NULL

curk
prevk

```

### Run the **Decision Engine** in the **post-basal** state (empty history):

Here, the `steps_factor` and `minutes_inc` are used to define the next
target.

```{r decide}
state0 <- list(
  history = list(),
  last_X = NA_integer_, last_Y = NA_integer_, last_Z = NA_integer_,
  consecutive_fails = 0L
)

res <- stepinWearable:::decide_message(
  state = state0, 
  cur_k = curk, 
  prev_k = prevk, 
  nombre = "Maria",
  steps_factor = 1.05,
  minutes_inc = 5,
  t = 0
)
str(res)
cat(res$text)
```

> In a real app session, these KPIs and the decision are calculated
> based on the **user-selected window** (and its previous window and
> target) and the participant's **persistent state**.

------------------------------------------------------------------------

## Result Saving Flow

When the user clicks **Save State** in the module, two actions occur:

1.  A `.RData` file is saved containing **only the selected window**
    (overwriting if the t-index file already exists):

    -   `steps_ts`: Minute-level data **filtered to the window**.

    -   `dsum`: Daily summary **of that window**.

    -   `medians`: List with medians from the window, including
        `median_steps_day`, cadence medians, and
        `window_start`/`window_end` dates.

2.  The participant state file (`*_state.rds`) is updated (upserted) by
    t-index with:

    -   The window's `kpis`.

    -   The message keys (`auto_message_key`, `final_message_key`).

    -   The new targets for the next cycle: `last_X`, `last_Y`,
        `last_Z`.

    -   The new `consecutive_fails` count.

> **Note:** If the **last 3 days of the window** are missing
> information, the targets are **not** updated in the state file.

------------------------------------------------------------------------

## Quick References

-   `?daily_summary` — Daily summary with total steps and minutes at
    ≥80/90/100/110/120.

-   `?select_windows` — Construction of the current and previous
    analysis windows.

-   `?kpis` — KPIs per window.

-   `?decide_message` — Template choice and adjustment of X/Y/Z targets.

-   `?mod_intervention_server` — Full module flow (automatic message →
    optional override → Save State).

------------------------------------------------------------------------

## Reproducibility

This vignette uses **synthetic data** generated at compile time and does
not include real participant data.
