---
title: "Fitbit Processing: Milestone Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitbit Processing: Milestone Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width = 7, fig.height = 4
)
```

## Overview

This vignette describes the files the **Intervention** module reads and writes:

-   **Participant registry** (`participant_info.rds`) – a simple ID ↔ name lookup used by the app.

-   **Per-participant state** (`state.rds`) – the intervention history and last targets for each participant.

-   **Milestone files** (`<ID>-<YYYYMMDD>-t<N>.RData`) – per-cycle saves that include:

    -   `steps_ts`: minute-level step counts for the selected 14-day window

        `dsum`: daily summary for that window.

    -   `medians`: precomputed medians for convenience.

### Where are these files?

By default the app uses directories derived from the package options set at load time (see `?stepinWearable::run_app`):

-   **Output root**: `getOption("stepin.output_dir")`

-   **Fitbit root**: `file.path(getOption("stepin.output_dir"), "output", "fitbit")`

-   **Participant folder**: `file.path(<fitbit root>, <participant_id>)`

The app creates folders on demand. You can override locations via R options in `.Rprofile`, e.g.:

```{r setup}
options(
  stepin.output_dir      = "~/stepinWearable-data",
  stepin.fitbit_download = "~/stepinWearable-data/download/fitbit"
)
```

> These defaults are safe on any machine and don’t require access to an institutional server.

## Participant registry: `participant_info.rds`

-   **Path:** `file.path(<fitbit root>, "participant_info.rds")`

-   **Type:** `list` with the following elements:

    -   **id:** participant-assigned ID

    -   **name:** participant-assigned name (only used for rendering the whatsapp message.

    -   **updated:** time stamp for the last time this file was saved.

## State file: `state.rds`

-   **Path:** `file.path(<fitbit root>, <participant_id>, "state.rds")`

-   **Type:** a named **list** with:

Top-level fields:

| name                | type    | meaning                                     |
|------------------|---------------|---------------------------------------|
| `history`           | list    | One entry per cycle (*t*-index). See below. |
| `last_X`            | numeric | Last target for daily steps (X).            |
| `last_Y`            | integer | Last target minutes/day at cadence Z (Y).   |
| `last_Z`            | integer | Last cadence threshold Z (80, 90, or 100).  |
| `consecutive_fails` | integer | Current streak of failed cycles.            |

Each element of `history` is a named list with:

| name | type | meaning |
|------------------------|------------------------|------------------------|
| `t_index` | integer | Cycle index *t*. |
| `date_str` | character | `YYYYMMDD` start date of the underlying minute series. |
| `start_date` | Date | First day of the 14-day analysis window. |
| `end_date` | Date | Last day of the 14-day analysis window. |
| `kpis` | list | Window KPIs (e.g., `med_steps_day`, `med_steps_80plus/90plus/100plus`). |
| `steps_factor` | numeric | Factor used to compute next X. |
| `minutes_inc` | integer | Minute increment applied to Y. |
| `target_steps` | numeric | Next X (steps/day). |
| `target_minutes` | integer | Next Y (minutes/day at Z). |
| `target_cadence` | integer | Next Z (80, 90, 100). |
| `auto_message_key` | character | Template key picked by the engine. |
| `final_message_key` | character | Template key actually used (after override, if any). |
| `manual_override` | logical | `TRUE` if final key differs from auto. |
| `message` | character | Rendered WhatsApp message stored for audit. |

## Milestone files: `<ID>-<YYYYMMDD>-t<N>.RData`

-   **Path pattern:** `file.path(<fitbit root>, <participant_id>, sprintf("%s-%s-t%d.RData", <ID>, <YYYYMMDD>, <t>))`

-   **Contains:** three objects:

### 1) `steps_ts` (minute level)

A `data.frame` (or `tibble`) with at least:

| column      | type        | meaning                                           |
|---------------|---------------|-------------------------------------------|
| `timestamp` | POSIXct     | Minute timestamp (local tz used for comparisons). |
| `steps`     | numeric/int | Step count for that minute.                       |

### 2) `dsum` (daily summary for the selected 14-day window)

A `data.frame` with:

| column          | type | meaning                           |
|-----------------|------|-----------------------------------|
| `date`          | Date | Day.                              |
| `steps_day`     | num  | Total steps that day.             |
| `steps_80plus`  | num  | Minutes/day at cadence ≥ 80 spm.  |
| `steps_90plus`  | num  | Minutes/day at cadence ≥ 90 spm.  |
| `steps_100plus` | num  | Minutes/day at cadence ≥ 100 spm. |

### 3) `medians` (precomputed convenience values)

A named list:

-   `window_start`, `window_end` (Date),

-   `median_steps_day`,

-   `median_steps_80plus`, `median_steps_90plus`, `median_steps_100plus`.

## Filenames and conventions (summary)

-   **Participant registry:**\
    `file.path(<output_dir>, "output", "fitbit", "participant_info.rds")`

-   **Participant folder:**\
    `file.path(<output_dir>, "output", "fitbit", <participant_id>)`

-   **State file:**\
    `file.path(<participant folder>, "state.rds")`

-   **Milestones:**\
    `file.path(<participant folder>, sprintf("%s-%s-t%d.RData", <ID>, <YYYYMMDD>, <t>))`

These conventions are enforced by the app when you press **Save Intervention State**.

## See also

-   The **Intervention** module UI/server help pages for how the app reads/writes these files.

-   The README for how to set default directories via `options()`.
